---
title: "Methods 4 -- Portfolio Assignment 2"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

-   *Type:* Group assignment
-   *Due:* 3 April 2022, 23:59

Hello CogSci's :)

In this portfolio, you are asked to do four tasks:

\- Make a DAG for something

\- Simulate data that fits the DAG

\- Use linear models to confirm that the DAG fits the data

\- Mess it up.

Each of the four tasks have some sub-steps.\
Report briefly what you find, for example in a markdown document, for
example called report.md so that the poor TA can easily get an overview
before looking in your code :)

Then you can also make a (brief!) explanation of the phenomenon you are
DAGGIN, simulating and modelling.

Looking forward!

```{r setup, include=FALSE}
pacman::p_load(usethis,dplyr, tidyverse, ggplot2, rstan, dagitty,ggdag)
library(rethinking)
```

## Task 1: The DAG

\- **Come up with an** incredibly interesting and scientifically
important made-up **example** for a phenomenon to investigate. Decide on
two variables (an outcome and a predictor) that you would like to
investigate the relation between. If in doubt, you **can be inspired by
Peter's amazing example** on the next page.

*In our example, we wish to investigate the effect of the predictor corona fear (CF) on*
*the outcome variable energy level (continuous).*

*Predictor: corona fear (continuous variable)* 
*Outcome variable: energy level (continuous variable), depends on germ level as germs gobble up nutrients and energy.*
*Sanitizing: continuous variable (count variable)* 
*Social exposure: a measure of how much social mingling you engage in (social circle size, social activities etc.)* 
*Germ level: the level of vira germs in your body. We have no biological evidence for a variable of this type, but for the sake of simplicity we assumes that the level of germs present in your body can be describes as a normally distributed continuous variable*

\- **Make a DAG** for the phenomenon. Make it medium complicated: that
means, make sure there are some different kinds of relations (see next
step). Change it if you don't get anything interesting for the next
steps.\
**Draw it** somehow (on paper, in R, laser engraved in diamond).\
**Code it** in dagitty (this is a nice tool:
<http://dagitty.net/dags.html> )


```{r}

# Twisted DAG corona

dag_corona <- dagitty( "dag {
    corona_fear -> sanitizing
    sanitizing-> germ_level
    OCD -> sanitizing
    OCD -> corona_fear
    social_exposure -> germ_level
    corona_fear -> social_exposure
    germ_level -> energylvl
}")

# Plotting the DAG
coordinates(dag_corona) <- list( 
  x=c(corona_fear =-1, OCD = -3, social_exposure =0, sanitizing =-2, germ_level=-1, energylvl= -1) , 
  y=c(corona_fear =-1, OCD = -1, social_exposure =0, sanitizing = 0, germ_level=1, energylvl=2) )

drawdag(dag_corona)

```

\- Find **elemental forms of variable relations** in the DAG (i.e.,
forks, pipes, colliders, and their descendants).

Pipe

(SE -\> GERM LEVEL -\> energylvl)

(Cf --\> SE --\> GERM)

(CF -\> SAN -\> GERM)

(S --\> GERM --\> energylvl)

(OCD --\> CF --\> SE)   

(OCD --\> SAN --\> GERM LEVEL)

Fork

(SAN \<- CF -\> SE)

(CF \<- OCD -\> SAN)

Collider

(SAN -\> GERM LEVEL \<- SE)

(CF-\> SAN \<- OCD)

Descendant 

(OCD --\> CF --\> SE, CF --\> SAN )

*We expect OCD to affect your corona fear in such way that a higher level of OCD increase the level of corona fear.* 

*A typical symptom of OCD is an abnormal fear of germs, which results in highly frequent use of sanitizer. Thus, we assume that the amount of sanitizing will increase with the level of OCD*.

*We expect the variable SE to depend on your corona fear, as we assume that a person who is really scared of COVID would have a much lower social exposure, whereas a person who didn't fear corona at all would have a correspondingly larger social exposure*-

*The same goes for the relation between corona fear and sanitizing. We expect a person with a high corona fear to have a much higher*
*daily frequency of sanitizing than a person with no corona fear at all*.

*We would expect germ level to depend on your social exposure, as a lot of social exposure implies more exposure to germs.* 

*The same goes for the relation between germ level and sanitizing, as*
*sanitizing works by killing most germs, preventing them from entering your body*

*Lastly, we expect the level of germs to determine your energy level. With a low level of germs,*
*the germs will not take up much energy from your body, but with a higher number of germs in your system, the germs will take more energy* 

\- Find out **what variables to include (and not include)** in a
multiple linear regression to avoid 'back door' (AKA non-causal) paths.
Do this first with your eyes and your mind. Then you can use dagitty's
function `adjustmentSets()`.

```{r}
# Shutting the backdoor - analyzing the graph to block the backdoor
adjustmentSets(dag_corona,exposure="corona_fear", outcome="energylvl")
```

OCD is a backdoor that we should control for (i.e., include in a our model) in order to find the full causal effect of corona fear on energylvl. 


\- Find out which **conditional independencies** the DAG implies. First
with the mind, then with daggity's function
`impliedConditionalIndependencies()`.


```{r}
# Deriving our DAG’s conditional independencies
impliedConditionalIndependencies(dag_corona)
```
   
  OCD _||_ enrg | grm_
  OCD _||_ enrg | sntz, scl_
  OCD _||_ enrg | crn_, sntz
  OCD _||_ grm_ | sntz, scl_
  OCD _||_ grm_ | crn_, sntz
  OCD _||_ scl_ | crn_
  crn_ _||_ enrg | grm_
  crn_ _||_ enrg | sntz, scl_
  crn_ _||_ grm_ | sntz, scl_
  enrg _||_ sntz | grm_
  enrg _||_ scl_ | grm_
  sntz _||_ scl_ | crn_

The above shows the testable implications of our DAG, which we will go through in task 3 to test if any of them can be falsified - i.e. to know if the DAG is compatible with the data. If any of the conditional independencies are not fulfilled, we can falsify our DAG.      


\- Find the full list of **Markov equivalent** DAGS. Use daggity's
function `equivalentGraphs()`.

The equivalentGraphs() function outputs a set of DAGs with the same
conditional independencies is known as a Markov equivalence set.

The list outputted allow show us 8 DAGS that would have the exact same
conditional independencies as our dag_corona has.

```{r}
equivalentDAGs(dag_corona)
```

## Task 2: The data

\- **Simulate some data that fits the DAG.** There are many ways to do
this. A simple way is just to sample one variable from a normal
distribution which has another variable as mean. McElreath does this in
the book a few times, and you can use this as inspiration.

```{r}

set.seed(1)
N = 100000

# Simulating variables:

# OCD
OCD <- sample(1:5, size=N, replace=TRUE, prob=c(.8, .10, .05, .03, .02))

# Corona fear (assuming that it is normally distributed and dependent on OCD)
corona_fear <-rnorm(N, mean=20+(10*OCD), sd=4) 
hist(corona_fear)

# Social exposure (is dependent on corona_fear)
social_exposure <- rnorm(N, mean= (35-(corona_fear/3)), sd= 1) #
plot(corona_fear,social_exposure)

# Sanitizing (is dependent on corona_fear and OCD)
sanitizing <-  rnorm(N, mean= (corona_fear-8)+(OCD*2), sd=2)
hist(sanitizing)
sanitizing_skrrt <-  rnorm(N, mean= (corona_fear-9), sd=2)
hist(sanitizing_skrrt)

# Germ level (is dependent on both sanitizing & whether you wear face mask or not)
germ_level <-  rnorm(N, mean= 5000 - ((sanitizing*10)*(1+social_exposure)), sd=100)

# Probability of being energylvl (depends on your germ level)
energylvl <-  rbinom(N, size= 1, prob= inv_logit(scale(germ_level)))
#energylvl <- rnorm(N, mean=-0.8*germ_level+4200, sd = 2)
plot(energylvl,germ_level)
hist(energylvl)
hist(germ_level)

plot(germ_level, inv_logit(scale(germ_level)), main= "Sigmoid plot: probability of being energylvl given germ level")

df <- tibble(germ_level, sanitizing, energylvl, social_exposure, corona_fear, OCD) 

# creating df
#df$energylvl <- ifelse(df$energylvl=="0",1,2)

ls.str(df)
```

*Standardizing variables*
Standardizing variables: OBS CHRIS: SHOULD WE STANDARDIZE A DICHOTOMOUS VARIABLE, SE AND OCD!!!!!! 
```{r}
df <- df %>% 
  mutate(germ_std=scale(germ_level),
         sanitizing_std= scale(sanitizing),
         energylvl_std=scale(energylvl),
         social_exposure_std=scale(social_exposure),
         corona_fear_std= scale(corona_fear),
         #OCD_std= scale(OCD)
         )

#ls.str(df)
```



## Task 3: Statistics
\- Run **multiple linear regression**s to **test the conditional
independencies** **implied by your DAG**. Make sure to avoid backdoor
paths. See that the linear model shows the conditional independencies
implied by your DAG, implying that the data and the DAG are compatible
(if the linear model doesn't show the conditional independencies implied
by the DAG, the data and the DAG doesn't fit).

### Testing the conditional independencies
Model 1: crn_ _||_ grm_ | fc_m, sntz. 
Testing whether the causal coefficient germ_level is insignificant when stratifying by social_exposure and sanitizing. 
```{r}

m1<- quap(
    alist(
        corona_fear_std ~ dnorm( mu , sigma ) ,
        mu <- a  + bGermLevel*germ_std + bFaceMask*social_exposure_std + bSanitizing*sanitizing_std,
        a ~ dnorm( 0 , 0.2 ) ,
        bGermLevel~ dnorm( 0 , 1 ),
        bFaceMask~ dnorm( 0 , 1 ) ,
        bSanitizing ~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m1),by.model=TRUE )

```
As seen in the above plot, corona fear and germ level becomes independent (i.e., the model with almost no uncertainty estimates the coefficient bGermLevel to be very close to zero ) when stratifying by social_exposure and sanitizing. 


Model 2: crn_ _||_ infc | grm_
Testing whether the causal coefficient infc (being energylvl or not) is insignificant when stratifying by germ_level. 
```{r}

m2<- quap(
    alist(
        corona_fear_std ~ dnorm( mu , sigma ) ,
        mu <- a + benergylvl*energylvl_std  + bGermLevel*germ_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl ~ dnorm( 0 , 1 ) ,
        bGermLevel~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m2), par=c("benergylvl", "bGermLevel" ))

```
As seen in the above plot, corona fear and energylvl becomes independent (i.e., the model confidently estimates the coefficient benergylvl to be very close to zero) when stratifying by germ_level. 


Model 3: crn_ _||_ infc_ | fc_m, sntz
Testing whether the causal coefficient benergylvl is insignificant when stratifying by social_exposure and sanitizing. 
```{r}

m3 <- quap(
    alist(
        corona_fear_std ~ dnorm( mu , sigma ) ,
        mu <- a + benergylvl * energylvl_std  + bFaceMask * social_exposure_std + bSanitizing * sanitizing_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl ~ dnorm( 0 , 1 ),
        bFaceMask ~ dnorm( 0, 1 ),
        bSanitizing~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 ) # Contains no more info than an average deviation. The avg. is just the inverse of the rate, 1. 
    ) , data = df )


plot(coeftab(m3),by.model=TRUE )
```
As seen in the above plot, benergylvl becomes independent (i.e., the model does with almost no uncertainty estimate the coefficient benergylvl to be very close to zero) when stratifying by social_exposure and sanitizing. 


Model 4: fc_m _||_ infc_ | grm_
Testing whether the causal coefficient infc (being energylvl or not) significantly close to zero when stratifying by germ_level in a model that has social_exposure as outcome variable.
```{R} 

m4 <- quap(
    alist(
        social_exposure_std ~ dnorm( mu , sigma ) ,
        mu <- a + benergylvl*energylvl_std  + bGermLevel*germ_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl ~ dnorm( 0 , 1 ) ,
        bGermLevel~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m4),by.model=TRUE)

```
As seen in the above plot, bGermLevel becomes independent (i.e., the model does with almost no uncertainty estimate the coefficient benergylvl to be very close to zero) when stratifying by germ level.

Model 5: fc_m _||_ sntz | crn_
Testing whether the causal coefficient of sanitizing becomes significantly close to zero when stratifying by corona fear in a model that has social_exposure as outcome variable.
```{R} 

m5 <- quap(
    alist(
        social_exposure ~ dnorm( mu , sigma ) ,
        mu <- a + bSanitizing*sanitizing_std  + bCoronaFear*corona_fear_std,
        a ~ dnorm( 0 , 0.2 ) ,
        bSanitizing ~ dnorm( 0 , 1 ) ,
        bCoronaFear~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m5),by.model=TRUE)
```
The coefficient bSanitizing is still close to zero, but not as confidently estimated as zero as in the previous plots. OBS CHRIS: can we say that the coefficient bSanitizing is close to zero (thus in accordance with our dag), or does such as small deviation from 0 count as a violation of the implications of our DAG???? 

Model 6: infc_ _||_ sntz | grm_
Testing whether the causal coefficient of sanitizing becomes significantly close to zero when stratifying by germ level in a model that has energylvl as outcome variable.
```{R} 

m6 <- quap(
    alist(
        energylvl_std ~ dnorm( mu , sigma ) ,
        mu <- a + bSanitizing*sanitizing_std  + bGermLevel*germ_std,
        a ~ dnorm(0,0.2) ,
        bSanitizing ~ dnorm(0,1) ,
        bGermLevel~ dnorm(0,1),
        sigma ~ dexp(1)
    ) , data = df )

plot(coeftab(m6),by.model=TRUE)

```
As seen in the above plot, bSanitizing becomes independent from infection (i.e., the model does with almost no uncertainty estimate the coefficient bSanitizing to be very close to zero) when stratifying by germ level.


Model 7: OCD _||_ fc_m | crn_
Testing whether the causal coefficient of social_exposure becomes significantly close to zero when stratifying by corona fear in a model that has OCD as outcome variable.
```{r}
m7 <- quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a + bFaceMask*social_exposure_std + bCoronaFear*corona_fear_std,
        a ~ dnorm( 0 , 0.2 ) ,
        bCoronaFear ~ dnorm( 0 , 0.5 ) ,
        bFaceMask ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 ) # OBS : why exp? It doesn't change anything changing it to dnorm or dunif
    ) , data = df )

plot(coeftab(m7),by.model=TRUE)

```
As seen in the above plot, bFaceMask becomes independent from OCD (i.e., the model does with almost no uncertainty estimate the coefficient bFaceMask to be very close to zero) when stratifying by corona fear.



Model 8: OCD _||_ grm_ | fc_m, sntz (OBS, not working)
Testing whether the causal coefficient of germ level becomes significantly close to zero when stratifying by face mask and sanitizing  in a model that has OCD as outcome variable.
```{r}
m8 <- quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a + bGermLevel*germ_std + bFaceMask*social_exposure_std + bSanitize*sanitizing_std ,
        a ~ dnorm( 0 , 0.2 ) ,
        bGermLevel ~ dnorm( 0 , 1 ) ,
        bFaceMask ~ dnorm( 0 , 1 ) ,
        bSanitize ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 ) # OBS : why exp? It doesn't change anything changing it to dnorm or dunif
    ) , data = df )

plot(coeftab(m8),by.model=TRUE)
precis(m8)
```
(OBS) Chris: coefficient is very low - but why not almost zero? Is it the model (should we have different slopes/intercepts dependent on our categorical predictor social_exposure) or does it have something to do with our data generation process???


Model 9: OCD || grm_ | crn_, sntz
Testing whether the causal coefficient of germ level becomes significantly close to zero when stratifying by corona fear and sanitizing in a model that has OCD as outcome variable.
```{r}
m9<- quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a  + bGermLevel*germ_std + bCoronaFear*corona_fear_std + bSanitizing*sanitizing_std,
        a ~ dnorm( 0 , 0.2 ) ,
        bGermLevel~ dnorm( 0 , 1 ) ,
        bCoronaFear~ dnorm( 0 , 1 ),
        bSanitizing ~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m9),by.model=TRUE)
```
As seen in the above plot, bGermLevel becomes independent from OCD (i.e., the model does with almost no uncertainty estimate the coefficient bGermLevel to be very close to zero) when stratifying by sanitizing and corona fear.



Model 10: OCD || infc | grm_
Testing whether the causal coefficient of energylvl becomes significantly close to zero when stratifying by germ level in a model that has OCD as outcome variable.
```{r}
m10<- quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a  + benergylvl*energylvl_std + bGermLevel*germ_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl ~ dnorm( 0 , 1 ),
        bGermLevel~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m10),by.model=TRUE)
```
As seen in the above plot, benergylvl becomes independent from OCD (i.e., the model does with almost no uncertainty estimate the coefficient benergylvl to be very close to zero) when stratifying by germ level. 



Model 11: OCD _||_ infc | fc_m, sntz
Testing whether the causal coefficient of energylvl becomes significantly close to zero when stratifying by face mask and sanitizing in a model that has OCD as outcome variable.
```{r}

m11 <-  quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a + benergylvl*energylvl_std  + bsocial_exposure*social_exposure_std + bSanitizing*sanitizing_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl~ dnorm( 0 , 1 ) ,
        bsocial_exposure~ dnorm( 0 , 1 ),
        bSanitizing~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m11),by.model=TRUE)

```
As seen in the above plot, benergylvl becomes independent from OCD (i.e., the model does with almost no uncertainty estimate the coefficient benergylvl to be very close to zero) when stratifying by face mask.


Model 12: OCD _||_ infc | crn_, sntz 
Testing whether the causal coefficient of energylvl becomes significantly close to zero when stratifying by corona fear and sanitizing in a model that has OCD as outcome variable.
```{r}
m12<- quap(
    alist(
        OCD ~ dnorm( mu , sigma ) ,
        mu <- a + benergylvl*energylvl_std  + bCorona_fear*corona_fear_std + bSanitizing*sanitizing_std,
        a ~ dnorm( 0 , 0.2 ) ,
        benergylvl~ dnorm( 0 , 1 ) ,
        bCorona_fear~ dnorm( 0 , 1 ),
        bSanitizing~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m12),by.model=TRUE)
```
As seen in the above plot, benergylvl becomes independent from OCD (i.e., the model does with almost no uncertainty estimate the coefficient benergylvl to be very close to zero) when stratifying by sanitizing and corona fear. 


## Task 4: Messing it up

\- Try and **deliberately have an open back door path** and see if you can get wrong inference.

*Comparing the two models*
*We get wrong inference if we dont include OCD in our model*
*Not including OCD in the model results in a different estimate of the level of corona fear's influence on the chance of whether you get energylvl or not, that is zero influence of corona fear on influence.*
*Contrarily, including OCD in the model and thus closing the backdoor, results in an estimate of a negative relationship between level of corona fear and the chance of getting energylvl.*


OBS comment

```{r}

m_backdoor<- quap(
    alist(
        energylvl ~ dnorm( mu , sigma ) , # (OBS ask Chris about standardizing dichotomous variable)
        mu <- a + bCorona_fear*corona_fear_std,
        a ~ dnorm( 0 , 0.2 ) ,
        bCorona_fear ~ dnorm( 0 , 1 ),
        sigma ~ dexp( 1 )
    ) , data = df )

plot(coeftab(m_backdoor),by.model=TRUE)
```

We assume that the logit transformation of the outcome variable (energylvl) has a linear relationship with the predictor variable (corona fear).

```{r}
backdoor <- quap(
    alist(
        energylvl ~ dbinom(2 , p) , # (OBS ask Chris about standardizing dichotomous variable)
        logit(p) <- a + bCorona_fear*corona_fear_std,# (OBS ask Chris about standardizing dichotomous variable)
        a ~ dnorm( 0 , 0.2 ) ,
        bCorona_fear ~ dnorm( 0 , 1 ),
        bOCD ~ dnorm( 0 , 1 )
    ) , data = df )

precis(backdoor)


### Inverse logit stuff
 # in the model WITH a backdoor, the estimated coef for the intercept is the log odds of a person with a corona fear of 0 being energylvl. In other words, the odds of being energylvl when the corona fear is 0 is exp(1.09)
exp(-.41)
inv_logit(-.41) # no backdoor, 0.3989121: one unit change in corona fear, means the log odds of energylvl (energylvl = 2) increases by 0.3989. 
inv_logit(-.36) # backdoor, 0.4109596

```

OBS comment
```{r}

nobackdoor<- quap(
    alist(
        energylvl ~ dbinom(2 , p) , # (OBS ask Chris about standardizing dichotomous variable)
        logit(p) <- a + bCorona_fear*corona_fear_std + bOCD*OCD,# (OBS ask Chris about standardizing dichotomous variable)
        a~ dnorm(0, 0.2) ,
        bCorona_fear~ dnorm( 0 , 1 ),
        bOCD~ dnorm(0 , 1 )
    ) , data = df)

precis(nobackdoor)


# Our estimated coefficient is the expected change in the log odds of being energylvl for a unit increase in the responding predictor variable holding the other predictor constant at a certain value. Each exponentiated coefficient is the ratio of two odds, or the change in odds in the multiplicative scale for a unit increase in the corresponding predictor variable holding other variables at certain value.  Here is an example.



```





\- Try and deliberately **simulate some data that doesn't fit the DAG**,
or **create a new DAG that doesn't fit the data**.


*We will now simulate some sata that does not fit the DAG*
```{r}
# Corona fear (assuming that it is normally distributed)
corona_fear_w <-rnorm(N, mean=50, sd=10)

# Face mask (depending on germ level)
social_exposure_w <- rbinom(N, size =1, prob= inv_logit(scale(germ_level)))

# Sanitizing (depending on germ_level)
sanitizing_w <- rnorm(N, mean= germ_level-20, sd=2)

# Germ level (depending on corona fear and face mask)
germ_level_w <-  rnorm(N, mean= 5000 - ((corona_fear*10)*(1+social_exposure)), sd=100)

# energylvl (not depending on any of the variables)
energylvl_w <-  rbinom(N, size= 1, prob=0.2)


# OCD level
OCD_w <- sample(1:5, size=N, replace=TRUE, prob=c(.8, .10, .05, .03, .02)) # 80% of not having OCD, apx. 2% chance for severe OCD

# creating df
df_w <- tibble(germ_level_w, sanitizing_w, energylvl_w, social_exposure_w, corona_fear_w, OCD_w) 
```





